### 2012-02-28 paper.tex
Further corrections, send to external (Søren Dahl) read-through

\twocolumn[
  \begin{@twocolumnfalse}
\noindent\LARGE{\textbf{Self-sustained carbon monoxide oxidation oscillations on size-selected platinum nanoparticles at atmospheric pressure$^\dag$}}
\vspace{0.6cm}

\noindent\large{\textbf{Robert Jensen\textit{$^{a}$}, Thomas Andersen\textit{$^{a}$}, Anders Nierhoff\textit{$^{a}$}, Thomas Pedersen\textit{$^{b}$}, Ole Hansen\textit{$^{b}$}, Søren Dahl\textit{$^{a}$} and Ib Chorkendorff\textit{$^{a}$}}}\vspace{0.5cm}
%Please note that \ast indicates the corresponding author(s) but no footnote text is required.

@@

\section{Experimental Methods}
The experiments were performed in Si-based $\mu$-reactors\cite{Henriksen2009}. The reactor design makes sure that all gases exposed to the catalyst under investigation is measured by the quadropole mass spectrometer (QMA) ensuring an extremely high sensitivity of the system.

@@

\section{Results}
An oscillation measurement is performed by mounting a sample in the setup directly after cluster deposition and anodic bonding. Initially a light-off CO-oxidation measurement is performed where the temperature is increased at 3\,$^\circ$C/min until the sample ignites and achieves full conversion of CO to CO$_2$. At O$_2$:CO ratio of $\sim0.08$ and a geometrical Pt coverage of $\sim0.1\%$ light-off occurs at approximately 180\,$^\circ$C. The temperature ramp is typically continued up to 260\,$^\circ$C whereafter the temperature is decreased at the same rate as the heating ramp. After the initial light-off the sample is cooled to room temperature. When at room temperature the temperature is again increased at 3\,$^\circ$C/min until sustained oscillations occur. A typical initial treatment is seen in Figure~\ref{fgr:initial_treatment} where both the symmetrical light-off ramp and the subsequent increase in temperature to achieve sustained oscillations is seen. Often the sample will show oscillations either directly upon light-off or on the falling temperature ramp.
\begin{figure}[h]
\centering
  \includegraphics[width=9cm]{initial_treatment.png}
  \caption{A typical example of the initial treatment of a new sample. After performing a light-off ramp where oscillations can been seen the temperature is increased to a constant value where sustained oscillations take place.}
  \label{fgr:initial_treatment}
\end{figure}
Nanoparticles with sizes ranging from 3\,nm to 9\,nm have been tested but 3\,nm particles with a geometrical coverage of 0.1\% of the reactor area gave the most stable oscillations. No consistent dependency on oscillation frequency and duty cycle was found that could be attributed to the size of the particles.

@@

An example of a single very slow oscillation is seen in Figure~\ref{fgr:full_oscillation}. After a short on-cycle (at $\sim$1030\,min.) where all CO is converted to CO$_2$ the reaction shows a fast deactivation. After the deactivation the sample quickly regains $\sim$30\% conversion (at $\sim$1040\,min.). Hereafter the conversion increases slowly over time until approximately 50\% conversion is reached (at $\sim$1118\,min.). Shortly after reaching 50\% conversion the sample again ignites and convert all CO to CO$_2$. After a short full conversion period the sample again deactivates and the cycle repeats itself.
\begin{figure*}
  \centering
  \includegraphics[width=17cm]{single_full_oscillation.png}
  \caption{A full oscillation period first showing a steep ignition of the sample followed by an almost immediate de-activation. For the next 65 minutes the sample slowly recovers activity until full conversion is reached again and the cycle repeats itself. This extremely long oscillation period is not commonly seen and was a result of carefull parameter tuning. Typical oscillation periods are normally between $\sim$30\,s and $\sim$30\,minutes.}
  \label{fgr:full_oscillation}
\end{figure*}

After prolonged measurements over several days an increase in oscillation period was observed as shown in Figure \ref{fgr:long_measurement}. Initially, the oscillation period is $\sim$2.5 minutes and increases linearly with time until 1500 minutes of total oscillation time. Hereafter the oscillations become more irregular as shown in Figure~\ref{fgr:long_measurement}. At experiment end the oscillation period is between 45 and 20 minutes.
\begin{figure}[h]
\centering
  \includegraphics[width=9cm]{summary_of_long_measurement.png}
  \caption{Summary of a sample oscillating under constant temperature, pressure and reactant composition. The sample went through a total of 439 oscillations in 4 days.}
  \label{fgr:long_measurement}
\end{figure}

@@

The oscillation frequency becomes more irregular with progressing experiment time as shown in Figure \ref{fgr:extracts}. However, the ratio between CO and CO$_2$ integrated over an full period remains constant throughout the entire experiment time, i.e. the samples produces the same amount of CO$_2$ independent of oscillation frequency.

@@

It has previously been shown using STM \cite{Hendriksen2002}, FT-IR \cite{Carlsson2006}, Monte Carlo simulations \cite{Zhdanov2002} as well as DFT \cite{Gong2004} that the activity of Pt towards CO oxidation at atmospheric pressures is highly dependent on the oxidation state of the surface. We are currently investigating new methods for gaining insight into the nature of the nanoparticles in the reactor but the low coverage and the design of the reactor makes this a challenging task. A detailed proposal for the reaction mechanism on Pd has been developed by Hendriksen \textit{et. al.}\cite{Hendriksen2010} and our data is consistent with this model.

According to the proposed model in ref.~\citenum{Hendriksen2010} the bare metal surface is less active than the oxidized surface. Oscillations originate from a switch between an oxidized surface and the bare metal surface. Initially, with CO present in the inlet gas a smooth metal surface is exposed due to the general high adsorption energy of CO on metal surfaces. As the Pt nanoparticles convert the CO to CO$_2$ the partial pressure of CO will decrease and a Pt oxide will start to form in the high partial pressures of O$_2$. After the oxidation of the nanoparticles they will roughen slowly during CO oxidation resulting in the formation of an increasingly rough oxide surface. As the oxide becomes rougher the bare metal surface is increasingly favored resulting in a return to the lower rate rough metal state. As the rate decreases the CO partial pressure increases which results in a transition from rough metal surface to a smooth metal surface. When the sample is sufficiently smooth it will again oxidize and thus complete the oscillation cycle.

In this study Pt nanoparticles were investigated which, compared to a single crystal or a thin film, is a very rough surface. It is to be expected that the particles will initially favor the reduced state. As the reaction runs, the conditions toward the outlet of the reactor will become more oxidizing and at the same time the reaction itself will smoothen the particles further pushing towards an oxidation of the particles. Gradually as the particles toward the outlet oxidizes the turn-over will increase and the general CO concentration will decrease promoting more and more particles to oxidize which will gradually be seen in the QMA as an increase in CO$_2$ and a decrease in CO. As the CO-concentration decreases the light-off temperature will decrease hence approaching the temperature of the sample resulting in a steep increase in reactivity. If not all particles are oxidized at this point they will oxidize much faster after light-off due to the low CO-concentration. The further increased rate will not be visible in the QMA since the reactor is already in full conversion. As most of the particles are now oxidized they will become more and more roughened and gradually return to the reduced state. Fewer and fewer oxidized particles will be responsible for the activity increasing the roughing of the remaining active particles. Just before the overall activity drops only a very small fraction of the particles actually participates in the reaction and the deactivation will happen very suddenly. Once these few particles eventually becomes sufficiently rough they will reduce and lose activity. It should be noted that our group has recently measured turnover frequencies of CO oxidation on Pt in excess of (Talværdier kommer lige om lidt...) ???/site/s corresponding to as little as $\sim$300 Pt nanoparticles contributing to the reactivity just before the end of the high-activity cycle.

The oscillation period increases with time and becomes increasingly less stable as shown in Figure~\ref{fgr:extracts}. This phenomena is consistent across all measured samples. The large scatter in oscillation period can be attributed to the smaller oscillations in between on/off cycles. Occasionally, these small amplitude oscillations will trigger a full switch thus introducing more short-period oscillations despite the trend of increasingly slower oscillation period. A possible explanation of the phenomenon could be sintering or roughening of the clusters. However, due to the fact that the overall conversion integrated over full oscillation periods is constant, we do not expect sintering to be cause, since this would also be expected to reduce the overall activity of the sample.

@@

Nanoparticles of several different sizes where tested which all showed oscillations. However, no change in duty cycle or frequency that could be attributed to the size of the nanoparticles was found. This may be due to the fact that even nominally identical samples also show a large variation in oscillation frequency and duty cycle thus possibly shadowing a nanoparticle size effect.

An interesting feature in the data is the very large temperature dependence. This property agrees well an earlier suggested model \cite{Hendriksen2010} and is, to our knowledge, the first time the temperature dependence of oscillations at atmospheric pressure has been measured.

>>>>

\twocolumn[
  \begin{@twocolumnfalse}
\noindent\LARGE{\textbf{Self-sustained carbon monoxide oxidation oscillations on size-selected platinum nanoparticles at atmospheric pressure}}
\vspace{0.6cm}

\noindent\large{\textbf{Robert Jensen, Thomas Andersen, Anders Nierhoff, Thomas Pedersen, Ole Hansen, Søren Dahl and Ib Chorkendorff}}\vspace{0.5cm}
%Please note that \ast indicates the corresponding author(s) but no footnote text is required.

@@

\section{Experimental Methods}
The experiments were performed in Si-based $\mu$-reactors\cite{Henriksen2009}. The reactor design makes sure that all gases exposed to the catalyst under investigation is measured by the quadropole mass spectrometer (QMS) ensuring an extremely high sensitivity of the system.

@@

\section{Results}
An oscillation measurement is performed by mounting a sample in the setup directly after cluster deposition and anodic bonding. Initially a light-off CO-oxidation measurement is performed where the temperature is increased at 3\,$^\circ$C/min until the sample ignites and achieves full conversion of CO to CO$_2$. At O$_2$:CO ratio of $\sim0.08$ and a geometrical Pt coverage of $\sim0.1\%$ light-off occurs at approximately 180\,$^\circ$C. The temperature ramp is typically continued up to 260\,$^\circ$C whereafter the temperature is decreased at the same rate as the heating ramp. After the initial light-off the sample is cooled to room temperature. When at room temperature the temperature is again increased at 3\,$^\circ$C/min until sustained oscillations occur. A typical initial treatment is seen in Figure~\ref{fgr:initial_treatment} where both the symmetrical light-off ramp and the subsequent increase in temperature to achieve sustained oscillations is seen. Often the sample will show oscillations either directly upon light-off or on the falling temperature ramp.
\begin{figure}[h]
\centering
  \includegraphics[width=9cm]{initial_treatment.png}
  \caption{A typical example of the initial treatment of a new sample. After performing a light-off ramp where oscillations can been seen the temperature is increased to a constant value where sustained oscillations take place. The non-zero value of CO during high conversion periods is consistent with the expected QMS background signals from CO$_2$ and O$_2$.}
  \label{fgr:initial_treatment}
\end{figure}
Nanoparticles with sizes ranging from 3\,nm to 9\,nm have been tested but 3\,nm particles with a geometrical coverage of 0.1\% of the reactor area gave the most stable oscillations. No consistent dependency on oscillation frequency and duty cycle was found that could be attributed to the size of the particles.

@@

An example of a single very slow oscillation is seen in Figure~\ref{fgr:full_oscillation}. After a short on-cycle (at $\sim$1030\,min.) where all CO is converted to CO$_2$ the reaction shows a fast deactivation. After the deactivation the sample quickly regains $\sim$30\% conversion (at $\sim$1040\,min.). Hereafter the conversion increases slowly over time until approximately 50\% conversion is reached (at $\sim$1118\,min.). Shortly after reaching 50\% conversion the sample again ignites and convert all CO to CO$_2$. After a short full conversion period the sample again deactivates and the cycle repeats itself.
\begin{figure*}
  \centering
  \includegraphics[width=17cm]{single_full_oscillation.png}
  \caption{A full oscillation period first showing a steep ignition of the sample followed by an almost immediate deactivation. For the next 65 minutes the sample slowly recovers activity until full conversion is reached again and the cycle repeats itself. This extremely long oscillation period is not commonly seen and was a result of careful parameter tuning. Typical oscillation periods are normally between $\sim$30\,s and $\sim$30\,minutes. Inserted circles illustrates the proposed model. At high conversion the platinum is oxidized (red) while in low conversion the sample is reduced (blue).}
  \label{fgr:full_oscillation}
\end{figure*}

After prolonged measurements over several days an increase in oscillation period was observed as shown in Figure \ref{fgr:long_measurement}. Initially, the oscillation period is $\sim$2.5 minutes and increases linearly with time until 1500 minutes of total oscillation time. Hereafter the oscillations become more irregular as shown in Figure~\ref{fgr:long_measurement}. At experiment end the oscillation period is between 45 and 20 minutes.
\begin{figure}[h]
\centering
  \includegraphics[width=9cm]{summary_of_long_measurement.png}
  \caption{Summary of a sample oscillating under constant temperature, pressure and reactant composition. The sample went through a total of 439 oscillations in 4 days. Time is defined from experiment start and hence includes initial treatment.}
  \label{fgr:long_measurement}
\end{figure}

@@

The oscillation frequency becomes more irregular with progressing experiment time as shown in Figure \ref{fgr:extracts}. However, the ratio between CO and CO$_2$ integrated over an full period remains constant throughout the entire experiment time, i.e. the samples maintains a constant overall rate independent of oscillation frequency.

@@

It has previously been shown using STM \cite{Hendriksen2002}, FT-IR \cite{Carlsson2006}, Monte Carlo simulations \cite{Zhdanov2002} as well as DFT \cite{Gong2004} that the activity of Pt towards CO oxidation at atmospheric pressures is highly dependent on the oxidation state of the surface. A detailed proposal for the reaction mechanism on Pd has been developed by Hendriksen \textit{et. al.}\cite{Hendriksen2010} and our data is consistent with this model.

According to the proposed model in ref.~\citenum{Hendriksen2010} the bare metal surface is less active than the oxidized surface. Oscillations originate from a switch between an oxidized surface and the bare metal surface. Initially, with CO present in the inlet gas a smooth metal surface is exposed due to the general high adsorption energy of CO on metal surfaces. As the Pt nanoparticles convert the CO to CO$_2$ the partial pressure of CO will decrease and a Pt oxide will start to form in the high partial pressures of O$_2$ thus increasing the rate. The model\cite{Hendriksen2010} suggests that the oxidized nanoparticles will roughen during CO oxidation resulting in the formation of an increasingly rough oxide surface. As the oxide becomes rougher the bare metal surface is increasingly favored resulting in a return to the lower rate rough metal state. As the rate decreases the CO partial pressure increases which results in a transition from rough metal surface to a smooth metal surface. When the sample is sufficiently smooth it will again oxidize and thus complete the oscillation cycle.

In this study Pt nanoparticles were investigated which, compared to a single crystal or a thin film, is a very rough surface. It is to be expected that the particles will initially favor the reduced state. As the reaction runs the gas composition towards the outlet of the reactor will become more oxidizing increasing the oxidation rate of the particles. Gradually, as the particles toward the outlet oxidizes the turn-over will increase and the general CO concentration will decrease promoting more and more particles to oxidize which will gradually be seen in the QMS as an increase in CO$_2$ and a decrease in CO. As the CO concentration decreases the light-off temperature will decrease hence approaching the constant temperature of the sample resulting in a steep increase in reactivity. It is important to note that the fraction of oxidized particles needed to achieve full conversion is not known. We are currently investigating new methods for gaining insight into the nature of the nanoparticles in the reactor but the low coverage and the design of the reactor makes this a challenging task. If not all particles are oxidized at light-off they will now oxidize much faster due to the low CO-concentration. The further increased rate will not be visible in the QMS since the reactor is already in full conversion. As most of the particles are now oxidized they will become more and more roughened and gradually return to the reduced state. Fewer and fewer oxidized particles will be responsible for the activity increasing the roughing rate of the remaining active particles. Just before the overall activity drops only a fraction of the particles participates in the reaction and the deactivation will happen very suddenly. Once these potentially few particles eventually becomes sufficiently rough they will reduce and lose activity hence completing the oscillation cycle. It should be noted that our group has recently measured turnover frequencies of CO oxidation on Pt in excess of $10^{6}$/site/s corresponding to as little as $\sim$300 Pt nanoparticles contributing to the reactivity just before the end of the high-activity cycle.

The oscillation period increases with time and becomes increasingly less stable as shown in Figure~\ref{fgr:extracts}. This phenomena is consistent across all measured samples. The large scatter in oscillation period can be attributed to the smaller oscillations in between on/off cycles. Occasionally, these small amplitude oscillations will trigger a full switch thus introducing more short-period oscillations despite the trend of increasingly slower oscillation period. A possible explanation of the phenomenon could be sintering or roughening of the clusters. However, due to the fact that the overall conversion integrated over full oscillation periods is constant we do not expect sintering to be cause since this would also be expected to reduce the overall activity of the sample.

@@

Nanoparticles of several different sizes were tested which all showed oscillations. However, no change in duty cycle or frequency that could be attributed to the size of the nanoparticles was found. This may be due to the fact that even nominally identical samples also show a large variation in oscillation frequency and duty cycle thus possibly shadowing a nanoparticle size effect.

An interesting feature in the data is the very large temperature dependence on oscillation frequency. This property agrees well an earlier suggested model \cite{Hendriksen2010} and is, to our knowledge, the first time the temperature dependence of oscillations at atmospheric pressure has been measured.

### 2012-02-28 supplementary.tex
Further corrections, send to external (Søren Dahl) read-through

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=9cm]{reactor.jpg}
  \caption{The Microreactor before the lid is anodicly bonded to the reactor}
  \label{fgr:reactor}
\end{figure}

@@

\section{Duty cycle}
Even though the period of the oscillations increases with time, the total integrated conversion doing a complete cycle is almost constant. In Figure~\ref{fgr:duty_cycles_supplemental}, the value of the integral of CO and CO$_2$ is plotted for every oscillation in the four-day long experiment. It is evident that the ratio between CO and CO$_2$ is almost constant and thus the overall activity of the sample is almost unchanged despite the development in the oscillation frequency.

>>>>

\renewcommand{\thefigure}{S\arabic{figure}}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=8cm]{reactor.jpg}
  \caption{The Microreactor before the lid is anodicly bonded to the reactor}
  \label{fgr:reactor}
\end{figure}

@@

\section{Duty cycle}
Even though the period of the oscillations increases with time, the total integrated conversion rate during a complete cycle is almost constant. In Figure~\ref{fgr:duty_cycles_supplemental}, the value of the mean value of CO and CO$_2$ is plotted for every oscillation in the four-day long experiment. It is evident that the ratio between CO and CO$_2$ is almost constant and thus the overall activity of the sample is almost unchanged despite the development in the oscillation frequency.
